Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-23 csf2/main.tex
Describe some syntax

\vspace{-1.5em}
\[
  \begin{array}{rcl}
    \src{\event[^b]} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \\
                     && \mid \src{\ev{Set \loc\ n}} \mid \src{\ev{Branch\ n}} \mid \src{\ev{Binop\ n}}\\
   && \src{\sandboxtag} := \src{\ctx} \mid \src{\comp} \hspace{3em}
   \src{\event} := \src{\event[^b];\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
    \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \\
                && \src{\library_{\ctx}} : \src{Vars} \to \src{\expr} \hspace{4em}
    \src{\library_{\comp}} : \src{Vars} \to \src{\expr} \\
  \end{array}
\]

$\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
%Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.

@@

\vspace{-1.5em}
\[
  \begin{array}{rcl}
    \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
  \end{array}
\]

@@

\vspace{-1.5em}
\[
  \begin{array}{rcl}
    \obj{D} &:=& \obj{ON} \mid \obj{OFF} \hspace{2em}
    \obj{\securitytag} := \obj{\lock} \mid \obj{\unlock} \\
    \obj{\event[^b]} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \mid \obj{Binop\ n} \mid \obj{Branch\ n} \\
    \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \mid \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
    \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \dots
  \end{array}
\]
\begin{center}
\footnotesize
  \typerule{$e-\obj{wrdoit}-\text{off}$}{
  }{
    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
  }{e-wrdoit-off}
\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
  \typerule{$e-\obj{get}-\in-$noleak}{
    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;m),\Delta_2} &
    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
    \obj{\sigma''}=\obj{\sigma\sqcap\sigma'}
  }{
    \exprevalo{\cfstate;\sandboxtag';m+1;\memstate}{x^\sigma[n^{\sigma'}]}{\cfstate;\sandboxtag';m+1;\memstate}{(H^{\sandboxtag}(\loc+n))^{\sigma''}}{{iGet}\ \loc\ n;\sandboxtag;\sigma''}
  }{to-e-get-in-noleak}
\end{center}

@@

\vspace{-1.5em}
\[
  \begin{array}{rcl}
    \ird{\event[^b]} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
    \ird{\expr[^b]} &:=& \dots \mid \ird{\lbarrier}
  \end{array}
\]

>>>>

$\src{L_{\tmssafe}}$ is a relatively naive language with second class functionns, the only values being numbers. 
It has simple binary operations, case analyses, C-like memory operations on arrays, as well as C-like function calls.
Programs have sets of pre-determined functions called libraries and it is assumed that there are functions part of some component or context.

\vspace{-1.5em}
\[
  \begin{array}{rrcl}
    \text{(Pre-Events)} & \src{\preevent} &:=& \src{\ev{Alloc \loc\ n}} \mid \src{\ev{Dealloc \loc}} \mid \src{\ev{Get \loc\ n}} \mid \src{\ev{Set \loc\ n}} \\
    \text{(Sandboxtag)} & \src{\sandboxtag} &:=& \src{\ctx} \mid \src{\comp} \\
    \text{(Events)} & \src{\event} &:=& \src{\preevent;\sandboxtag} \mid \src{\emptyevent} \mid \src{\lightning} \\ 
    \text{(Expressions)} & \src{\expr} &:=& \src{x} \mid \src{n} \mid \src{\lbinop{\expr[_1]}{\expr[_2]}} \mid \src{\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}} \\ 
                         &&&\mid \src{\llet{x}{\expr[_1]}{\expr[_2]}} \mid \src{\llet{x}{new\ \expr[_1]}{\expr[_2]}} \\
                         &&&\mid \src{\ldelete{x}} \mid \src{\lget{x}{\expr}} \mid \src{\lset{x}{\expr[_1]}{\expr[]}} \mid \src{stuck}\\
                         &&&\mid \src{\lcall{f}{\expr}} \mid \src{\lreturn{\expr}} \\
    \text{(Libraries)} &&& \src{\library_{\ctx}} : \src{Vars} \to \src{\expr} \hspace{2em}
    \src{\library_{\comp}} : \src{Vars} \to \src{\expr} \\
  \end{array}
\]

Crucially, $\src{L_{\tmssafe}}$ is statically typed, where the type system is inspired by $L^{3}$~\cite{morrisett2005L3,scherer2018fabulous}.
The type system of $\src{L_{\tmssafe}}$ enforces that every well-typed $\src{L_{\tmssafe}}$ program satisfies \gls*{tms} (\Cref{thm:wt:tms}).
%Hereby, the map $\thelocmap(\src{\loc})=\loc$ maps an $\Ltms$ location to a location $\loc$ of the specification trace model.

@@

\vspace{-1.5em}
\[
  \begin{array}{rrcl}
    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
  \end{array}
\]

@@

\vspace{-1.5em}
\[
  \begin{array}{rrcl}
    \text{(Mode Values)} & \obj{D} &:=& \obj{ON} \mid \obj{OFF} \\
    \text{(Security Tags)} & \obj{\securitytag} &:=& \obj{\lock} \mid \obj{\unlock} \\
    \text{(Pre-Event)} & \obj{\preevent} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \\
                       &&& \mid\ \obj{Binop\ n} \mid \obj{Branch\ n} \\
    \text{(Event)} & \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \obj{\emptyevent} \mid \obj{\lightning} \\
    \text{(Expressions)} & \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \\
                         &&&\mid\ \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
  \end{array}
\]
\begin{center}
\footnotesize
  \typerule{$e-\obj{wrdoit}-\text{off}$}{
  }{
    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
  }{e-wrdoit-off}
\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
  \typerule{$e-\obj{get}-\in-$noleak}{
    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;m),\Delta_2} &
    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
    \obj{\sigma''}=\obj{\sigma\sqcap\sigma'}
  }{
    \exprevalo{\cfstate;\sandboxtag';m+1;\memstate}{x^\sigma[n^{\sigma'}]}{\cfstate;\sandboxtag';m+1;\memstate}{(H^{\sandboxtag}(\loc+n))^{\sigma''}}{{iGet}\ \loc\ n;\sandboxtag;\sigma''}
  }{to-e-get-in-noleak}
\end{center}

@@

\vspace{-1.5em}
\[
  \begin{array}{rrcl}
    \text{(Pre-Events)} & \ird{\preevent} &:=& \dots \mid \ird{Spec} \mid \ird{Rlb} \mid \ird{Barrier}\\
    \text{(Expressions)} & \ird{\expr} &:=& \dots \mid \ird{\lbarrier}
  \end{array}
\]
